\chapter{Lorentzian geometry}

In this first chapter we introduce Lorentzian manifolds of constant curvature and observe that, as in the Riemannian case, two manifolds of constant sectional curvature $K$ are locally isometric. We start by recalling some basic definitions of Lorentzian geometry to set notation. We will then define what we mean by a manifold with maximal isometry group as spaces with such property provide \textit{models} of manifold with constant curvature: if we have $M$ a Lorentzian manifold with constant sectional curvature $K$ and maximal isometry group, then any Lorentzian manifold with constant sectional curvature $K$ carries a natural (Isom($M$), $M$)-atlas made of local isometries. \\
Simply connected spaces have maximal isometry group, but in general the converse is false. In particular, in the course of the dissertation, we will focus on $K=-1$ and in that case it will be convenient to use non-simply connected models. 

\subsection{Basic Definitions}
By a \textit{Lorentzian metric} on a $n+1$ manifold we mean a non degenerate $2$-tensor $g$ of signature $(n,1)$. A \textit{Lorentzian manifold} is a connected manifold $M$ equipped with a Lorentzian metric $g$.\\
In a Lorentzian manifold $M$ we say that a non-zero vector $v \in TM$ is \textit{time-like} if $g(v,v)<0$, \textit{space-like} if $g(v,v)>0$ and \textit{light-like} if $g(v,v)=0$. More generally, we say that a linear subspace $V \subset T_{x}M$ is \textit{spacelike, lightlike, timelike} if the restriction of $g$ to $V$ is positive definite, degenerate or negative definite respectively.\\
The set of lightlike vectors, together with the null vector, disconnects $T_{x}M$ into 3 regions: two convex open cones formed by timelike vectors, one opposite to the other, and the region of spacelike vectors. 

\textcolor{red}{Inserire figura del cono di Minkowski?}

\noindent As a consequence the set of timelike vectors in the total space $TM$ is either connected or is made by two connected components. In the latter case $M$ is said \textit{time-orientable}, and a time orientation is the choice of one of those components. Vectors in the chosen component are said \textit{future-directed}, vectors in the other component are said \textit{past-directed}. A spacetime is a Lorentzian manifold $(M,g)$ with the choice of a time orientation.\\
A differentiable curve is said \textit{timelike, spacelike, lightlike} if its tangent vector at every point is timelike, spacelike, lightlike respectively. The curve is said to be \textit{causal} if the tangent vector is either timelike or lightlike. \\ Given a point $x$ in a time-oriented Lorentzian manifold $M$, the \textit{future} of $x$ is the set $I^+(x)$ of points which are connected to $x$ by a future-directed causal curve. The \textit{past} of $x$, $I^-(x)$, is defined in an analogous way for past-directed causal curves \textcolor{red}{If we want to give a different pov on Domain of Dependence maybe we should introduce here the Alexander Topology}. \\As in the Riemannian setting, on a Lorentzian manifold $M$ there is a unique linear connection $\nabla$ which is symmetric and compatible with the Lorentzian metric $g$. We refer to it as the \textit{Levi-Civita connection} of $(M,g)$.\\ Following the analogy with Riemannian geometry, the Levi-Civita connection determines the Riemann curvature tensor defined by: 
\[
    R(u,v)w=\nabla_u\nabla_v w-\nabla_v\nabla_u w-\nabla_{[u,v]}w.
\]  

\noindent We then say that a Lorentzian manifold has constant sectional curvature $K$ if: 
\begin{equation}\label{sectionalcurvature}
    g(R(u,v)v,u)=K(g(u,u)g(v,v)-g(u,v)^2) 
\end{equation}
    


for every pair of vectors $u,v \in T_{x}M$ and every $x\in M$. Even tough the definition is analogous to the one given in the Riemannian realm, we recall that in the Lorentzian setting the sectional curvature can be defined only for planes in $T_{x}M$ where $g$ is non-degenerate. The manifold $M$ is said \textit{geodesically complete} if every geodesic is defined for all times, equivalently the exponential map is defined everywhere.\\

\subsection{Maximal isometry group and geodesic completeness} Two Riemannian manifolds $M$ and $N$ of constant curvature $K$ are locally isometric, the same statement holds for Lorentzian manifolds. The proof is analogous to the one in the Riemannian setting, and it is based on the use of the Cartan-Ambrose-Hicks theorem (the semi-Riemannian case is treated in \cite{piccione2005single}). More precisely we have the following:

\begin{lemma}\label{isometrie} Let $M$ and $N$ be Lorentzian manifolds of constant curvature $K$. Then every linear isometry $L:T_{x}M\to T_yN$ extends to an isometry $f:U\to V$ where $U, V$ are open neighbourhoods of $x, y$ respectively. Any two extensions $f:U\to V$ and $f^{\prime}:U^{\prime} \to V^{\prime} $ of $L$ coincide on $U\cap U^{\prime}. $ Moreover $L$ extends to a local isometry $f:M\to N$ provided that $M$ is simply connected and $N$ is geodesically complete. 
\end{lemma}

As a direct consequence of the aforementioned lemma we have: 
\begin{corollary}\label{122}
    Let $M$ and $N$ be simply connected, geodesically complete Lorentzian manifolds of constant curvature $K$. Then any linear isometry $L:T_xM\to T_yN$ extends to a global isometry $f:M\to N.$   
\end{corollary}

In particular, there is a unique simply connected geodesically complete Lorentzian manifold of constant curvature $K$ up to isometries. For instance for $K=0$ a model is the Minkowski space $\R^{n,1}$, that is $\R^{n+1}$ provided with the standard metric
\[
    g=dx_1^{2}+\dots+dx_n^{2}-dx_{n+1}^2.   
\]

Another consequence of Lemma \ref{isometrie} is that, fixing a point $x_0 \in M$, the set of isometries of $M$, which we will denote by Isom($M$), can be realized as a subset of ISO($T_{x_0}M, TM$) namely the fiber bundle over $M$ whose fiber over $x\in M$ is the space of linear isometries of $T_{x_0}M$ into $T_{x}M$. Now Isom($M$) considered with the operation of composition is a Lie group and the inclusion $\text{Isom}(M)\to \text{ISO}(T_{x_0}M, TM)$, is a differentiable proper embedding (\cite{okubo_1970}, Theorem 4.1). It follows that the maximal dimension of $\text{Isom}(M)$ is $\text{dim}(O(n,1))+n+1=(n+1)(n+2)/2$

\begin{definition}
    A Lorentzian manifold $M$ has \textit{maximal isometry group} if the action of Isom($M$) is transitive and, for every point $x \in M$, every linear isometry $L:T_{x}M\to T_xM$ extends to an isometry of $M$. 
\end{definition}

Equivalently $M$ has maximal isometry group if the above inclusion of Isom($M$) into ISO($T_{x_0}M, TM$) is a bijection. Hence if $M$ has maximal isometry group, its isometry group has maximal dimension, as a justification for the nomenclature used. 
From Corollary \ref{122}, every simply connected Lorentzian manifold $M$ has maximal isometry group if it has constant sectional curvature and is geodesically complete. The converse holds even without the simple connectedness assumption. Namely: 

\begin{lemma}\label{maximalisometry}
    If $M$ is a Lorentzian manifold with maximal isometry group, then $M$ has constant sectional curvature and is geodesically complete.
\end{lemma}

\begin{proof}
    Fix a point $x\in M$. As the identity component of $O(T_x M)\simeq O(n,1)$ acts transitively on spacelike planes, there exists a constant $K$ such that the equation \ref{sectionalcurvature} holds for every pair $(u,v)$ of vectors tangent at $x$ which generate a spacelike plane. Now, for every point $x\in M$ both sides of equations \ref{sectionalcurvature} are polynomial in $u,v \in T_xM$. Since the set of pairs $(u,v)$ which generate spacelike planes is open in $T_{x}M\times T_{x}M$, the equation \ref{sectionalcurvature} must hold for every pair of vectors $u,v \in T_xM$.\\ Since Isom($M$) acts transitively on $M$, it follows that $M$ has constant sectional curvature $K$.\\
    Let us now show that the manifold is geodesically complete. Suppose $\gamma$ is a parametrized geodesic with $\gamma(0)=x$ and $\gamma^{\prime} (0)=v\in T_xM,$ which is defined for a finite maximal time $T>0.$ Let $T_0=T-\epsilon>0.$ By assumption one can find an isometry $f:M\to M$ such that $f(x)=\gamma(T_0)$ and $df_x(v)=\gamma^{\prime}(T_0).$ Then $t\to f\circ\gamma(t-T_0)$ is a parametrized geodesic which provides a continuation of $\gamma$ beyond $T$, leading us to a contradiction. 
\end{proof}

We conclude these preliminaries on Lorentzian geometry with a result of classification which will be useful in the following, furthermore explaining our interest in spaces with maximal isometry group.
\begin{proposition}\label{classification}
    Let $M_K$ be a simply connected Lorentzian manifold of constant sectional curvature $K$ with maximal isometry group, and let $M$ be a Lorentzian manifold of constant sectional curvature $K$. Then: 
    \begin{itemize}
        \item M is geodesically complete if and only if there is a local isometry $p:M_K\to M$ which is a universal covering.
        \item $M$ has maximal isometry group if and only if $\text(Aut)(p:M_K\to M)$ is normal in Isom($M_K$). 
    \end{itemize}
\end{proposition}

\begin{proof}
    Suppose $M$ is geodesically complete, then by lifting the metric to the universal cover $\widetilde{M}$ one gets a simply connected geodesically complete Lorentzian manifold of constant sectional curvature $K$ which by Corollary \ref{122} is isometric to $M_K$. The covering map $p:M_K\to M$ is then a local isometry by construction. The converse is straightforward. \\   
    Now let $\Gamma=\text{Aut}(p:M_K\to M),$ which is a discrete subgroup of Isom($M_K$). Thus $M$ is obtained as the quotient $M=M_{K}/\Gamma$, where $\Gamma$ acts freely and properly discontinuously on $M_K$. The isometry group of $M$ is isomorphic to $N(\Gamma)/\Gamma$, where by $N(\Gamma)$ we denote the normalizer of $\Gamma$ in Isom($M_K$). The isomorphism is based on the observation that any isometry of $\widetilde{M}$ which normalizes $\Gamma$ descends to an isometry of $M$, and conversely the lifting of any isometry of $M$ must be in $N(\Gamma).$\\
    Hence the condition that $M$ has maximal isometry group is equivalent to the condition that every element $f$ of $\text{Isom}(M_K)$ descends in the quotient to an isometry of $M$. This is in turn equivalent to the condition that $f\Gamma f^{-1}=\Gamma$ for every $f\in \text{Isom}(M_K)$, which is the same as saying that $\Gamma$ is normal in $\text{Isom}(M_K)$.
\end{proof}

\begin{observation}
Since $\Gamma=\text{Aut}(p:M_K\to M)$ is discrete, being normal in $\text{Isom}(M_K)$ implies that elements of $\Gamma$ commute with the elements of the identity components of $\text{Isom}(M_K)$. This remark suggests that there are usually not many Lorentzian manifolds of constant sectional curvature with maximal isometry group. \textcolor{red}{probabilmente toglierei, abbastanza vaga come affermazione e mai utilizzata di fatto.}
\end{observation}

Finally, any isometry between connected open subsets of a Lorentzian manifold $M$ with maximal isometry group extends to a global isometry. In particular if $M_K$ is a Lorentzian manifold of constant sectional curvature $K$ with maximal isometry group, then any Lorentzian manifold $M$ of constant sectional curvature $K$ admits a natural $(\text{Isom}(M_K), M_K)$-structure whose charts are isometries between open subsets of $M$ and open subsets of $M_K$. We will refer to Lorentzian manifolds of constant sectional curvature $K$ with maximal isometry group as \textit{models} of constant sectional curvature $K$ (following Klein's terminology).  